---
title: "apply.Rmd"
author: "Damian Sanchez"
date: "4/4/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##
```{r}
?apply
?mapply
```
We `apply` functions allows us to apply a function to a vector ir list of values iteratively.
This helps minimize errors in code and makes the analyses more efficient.

With lapply() and apply() functions, we can only provide one argument to iterate on.

`sapply()` function simplifies the output to a vector (or the simplest data sturcture as possible) while `lapply()` returns an output in the form of a list.

With `mapply()`, we can provide multiple arguments. 


## Installing Library
```{r}
library(dplyr)
```

## Exercise 1 
```{r}
mass_from_length_theropoda <- function(length) {
  mass <- 0.73 * length^3.63
  return(mass)
}
theropoda_lengths <- c(17.8013631070471, 20.3764452071665, 14.0743486294308, 25.65782386974, 26.0952008049675, 20.3111541103134, 17.5663244372533, 11.2563431277577, 20.081903202614, 18.6071626441984, 18.0991894513166, 23.0659685685892, 20.5798853467837, 25.6179254233558, 24.3714331573996, 26.2847248252537, 25.4753783544473, 20.4642089867304, 16.0738256364701, 20.3494171706583, 19.854399305869, 17.7889814608919, 14.8016421998303, 19.6840911485379, 19.4685885050906, 24.4807784966691, 13.3359960054899, 21.5065994598917, 18.4640304608411, 19.5861532398676, 27.084751999756, 18.9609366301798, 22.4829168046521, 11.7325716149514, 18.3758846100456, 15.537504851634, 13.4848751773738, 7.68561192214935, 25.5963348603783, 16.588285389794)
theropod_masses <- mass_from_length_theropoda(length = theropoda_lengths)

my_list <- list(theropod_masses)
second_list <- c(my_list, list(c("Luna", "Avi", "Anita")))

second_list[[1]]

data.frame(theropod_masses, c("Anita", "Avi", "Luna", "Maria"))
```

```{r}
mass_from_length <- function(length, a = 0.73, b = 3.63) {
  mass <- a * length^b
  return(mass)
}
a_values <- c(0.759, 0.751, 0.74, 0.746, 0.759, 0.751, 0.749, 0.751, 0.738, 0.768, 0.736, 0.749, 0.746, 0.744, 0.749, 0.751, 0.744, 0.754, 0.774, 0.751, 0.763, 0.749, 0.741, 0.754, 0.746, 0.755, 0.764, 0.758, 0.76, 0.748, 0.745, 0.756, 0.739, 0.733, 0.757, 0.747, 0.741, 0.752, 0.752, 0.748)

b_values <- c(3.627, 3.633, 3.626, 3.633, 3.627, 3.629, 3.632, 3.628, 3.633, 3.627, 3.621, 3.63, 3.631, 3.632, 3.628, 3.626, 3.639, 3.626, 3.635, 3.629, 3.642, 3.632, 3.633, 3.629, 3.62, 3.619, 3.638, 3.627, 3.621, 3.628, 3.628, 3.635, 3.624, 3.621, 3.621, 3.632, 3.627, 3.624, 3.634, 3.621)

new_masses <- mass_from_length(length = theropoda_lengths)
# rm(new_lengths) The rm function allows us to remove objects from the R environment.

theropod_masses == new_masses
all(theropod_masses = new_masses) # tests that all values in the logical vector are equal to TRUE
all.equal(theropod_masses, new_masses)

mass_from_length(length = theropoda_lengths, a = a_values, b = b_values)

```
# Basic functions can naturally iterate through values in a vector to perform calculations.




```{r}
dino_data <- data.frame(theropoda_lengths, a_values, b_values)
dino_data <- dino_data %>%
  mutate(masses = mass_from_length(theropoda_lengths, a = a_values, b = b_values))

print(dino_data)
```
 `<<-` the scope operator or double arrow, allows creating and modifying variables in parent variables
 
 
## Exercise 2
```{r}
mass_from_length_max <- function(length) {
  if (length < 20) {
    mass <- 0.73 * length^3.63
    return(mass)
  } else {
    return(NA)
  }
}
masses_max <- sapply(theropoda_lengths, mass_from_length_max)
print(masses_max)
```

## Exercise 3
```{r}
dino <- read.csv(file = "../Downloads/dinosaur_lengths.csv")
get_mass_from_length_by_name <- function(lengths, species) {
  if (species == "Stegosauria") {
    a <- 10.95
    b <- 2.64
  } else if (species == "Theropoda") {
    a <- 0.73
    b <- 3.63
  } else if (species == "Sauropoda") {
    a <- 214.44
    b <- 1.46
  } else {
    a <- NA
    b <- NA
  }
  mass <- a*lengths^b
  return(mass)
}

masses <- mapply(get_mass_from_length_by_name, lengths = dino$lengths, species = dino$species)
dino %>%
  rowwise() %>%
  mutate(masses = get_mass_from_length_by_name(lengths, species)) %>%
  head(10)

library(ggplot2)
dino %>%
ggplot() +
  geom_histogram(mapping = aes(x=masses, color= species)) +
  facet_wrap(~species)
```

