---
title: "functions-after.Rmd"
author: "Damian Sanchez"
date: "3/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installing library package
```{r}
library(dplyr)
library(ggplot2)
```

#reading in the csv files 
```{r}
surveys <- read.csv(file = "../data-raw/surveys.csv")
species <- read.csv(file = "../data-raw/species.csv")
plots <- read.csv(file = "../data-raw/species.csv")
```

# combining surveys and species tables into single data frame
```{r}
combined_df <- inner_join(surveys, species, by ="species_id")
```

## writing a function that takes three arguments 
```{r}
get_species_count_by_year <- function(df, genus, species){
  species_df <- df %>%
  filter(genus == "Dipodomys" | species == "merriami")
  count_df <- species_df %>%
  group_by(year) %>%
  summarize(count = n())
return(count_df)
}
```

# using function to get the data frame of time series by year for the genus "Dipodomys" and the species "merriami" and make a graph
```{r}
count_df <- get_species_count_by_year(combined_df, "Dipodomys", "merriami")

ggplot(count_df, aes(x = year, y = count)) +
  geom_point() +
  labs(x = "Year", y = "Count", title = "Number of Dipodomys merriami individuals by year")
```
# Using the function to get the data frame series by year for the genus Chaetodipus and species penicillatus and make a graph

```{r}
count_df <- get_species_count_by_year(combined_df, "Chaetodipus", "penicillatus") 
ggplot(count_df, aes(x = year, y = count)) +
  geom_line(color = "blue", size = 2) +
  geom_point(color = "blue", size = 1) +
  labs(x = "Year", y = "Number of Individuals", title = "Number of Chaetodipus penicillatus individuals by year")
```

