---
title: "functions.Rmd"
author: "Damian Sanchez"
date: "3/21/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Exercise 1
```{r}
convert_pounds_to_grams <- function(pounds) {
     grams = 453.6 * pounds
     return(grams)
convert_pounds_to_grams(3.75)
 }
```
a) convert_pounds_to_grams is a function of pounds converted into grams 
b)
c)

## Using and Modifying functions
```{r}
get_mass_from_length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  return(mass)
get_mass_from_length_theropoda(16)  
}
```

## Create new version of this function
```{r}
get_mass_from_length <- function(length){
  mass <- 214.44 * length ^ 1.46
  return(mass)
get_mass_from_length(26)
}

```

## Introductory Section 
Functions allow for the creation of modular code, which makes it easier to manage. Functions can be reused in different parts of a data science project. Functions provide a way to abstract away the details of a complex algorithm or process.


# Exercise 3
```{r}
get_mass_from_length <- function(length){
  mass <- 214.44 * length ^ 1.46
  return(mass)
get_mass_from_length(22)
}
get_mass_from_length <- function(length){
  mass <- 39.9 * length ^ 2.6
  return(mass)
get_mass_from_length(16)
}
```

#  Exercise 4
```{r}
convert_kg_to_pounds <- function(kg){
  pounds = 2.205 * kg
  return(pounds)
  convert_kg_to_pounds(50)


get_mass_from_length <- function(length, a=10.95, b=2.64)
    mass = a * length ** b
    return(mass)
    get_mass_from_length(12)
}
```



