---
title: "wrangling-pipes.Rmd"
author: "Damian Sanchez"
date: "3/7/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The classic way of running code

For example, I want the square root of the mean of a sequence of numbers 

## Nested code
```{r}
numbers <- 1:300
mean(numbers)
sqrt(mean(numbers))
```

## Sequential code
In this case we create intermediate variables
```{r}
numbers <- 300:546
numbers <- 1:300
numbers_mean <- mean(numbers)
sqrt(x = numbers_mean)
```

## Piping code 
it can be implemented in R using the package `magrittr` It is a dependency of `diplyr`, so it is installed along.
```{r}
library(magrittr)
library(dplyr)
```

The original symbol of the pipe is `%>%`. 
But we also have a new symbol that is similar to bash `|>`.
The purpose of this is to eliminate the need of intermediate variables.
For the mean example

```{r}
1:300 %>% mean() %>% sqrt()
```

## Pipes with the surveys data set

```{r}
surveys <- read.csv(file = "../data-raw/surveys.csv")
str(surveys)
```


Calculate the mean of the year column using pipes 
```{r}
surveys$year %>% mean()
```

Calculate the mean of the weight column using pipes 
```{r}
surveys$weight %>% mean(na.rm = TRUE)
```

## Exercise 1
1)
```{r}
surveys <- read.csv(file = "../data-raw/surveys.csv")
```

2) Use select() to create a new data frame object called surveys1 with just the year, month, day. and species_id columns in that order.
```{r}
surveys1 <- select(surveys, year, month, day, species_id)
str(surveys1)
```

3) Create a new data frame called surveys2 with the year, species_id, and weight in kilograms of each individual, with no null weights. Use mutate(), select(), filter() with !is.na().
```{r}
surveys2 <- select(surveys, year, species_id, weight)
str(surveys2)
```

```{r}
surveys2 <- mutate(surveys2, weight_kg = weight/1000)
str(surveys2)
```

```{r}
surveys2 <- filter(surveys2,!is.na(weight_kg))
str(surveys2)
```

```{r}
surveys2 <- select(surveys2, year, species_id, weight_kg)
colnames(surveys2)
## surveys2[c(1,3)]
str(surveys)
## surveys2[ , c(1,3)]
## surveys2[ , c("year", "weight_kg")]
```



4. Use the filter() function to get all of the rows 
```{r}
surveys2_filtered <- filter(surveys2, species_id == "SH")
## surveys2_filtered
str(surveys2_filtered)
```

Exercise 2

```{r}
read.csv(file = "../data-raw/surveys.csv") %>% select( year, month, day, species_id) -> surveys1
str(surveys1) # Read the data set then select columns and assign to object
surveys %>% select( year, species_id, weight) %>% 
  mutate(weight_kg = weight/1000) %>% 
  filter(!is.na(weight)) %>%
  select(year, species_id, weight_kg) %>%
  filter(species_id == "SH") -> surveys_final
str(surveys_final)
```

## Exercise 3

```{r}

```


