---
title: "joining-vectors.Rmd"
author: "Damian Sanchez"
date: "3/16/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What are vectors?
- they are unidimensional matrices
- they can only hold one type of data, either numerical (integer or or double)
, character, or logical (complex numbers)

## What are data frames?
- they are two dimensional matrix, rows and columns
- can hold any type of data
- can only hold different types of data in a certain way:
- only column can have different data types
- within a column, all rows must have the same data type

- A data frame can also be defined as a collection of vectors that can be of different or the same type, vectors with all of the same length




```{r, error= TRUE}
str(surveys)
```
## Creating vectors-review 

A character vector with the function `c()`
```{r}
c("luna", "Avi", "Anita", "James", "Charles", "Damian", "Davinder") -> our_names
str(our_names)
1:7 # the colon operator creates a vector of numbers
1:7 -> my_numbers
-100:200
```

In how many steps or units does the colon operator increase? It increases in a unit of 1.

What do we do if I want to create a numeric sequence that increases in steps different than 1?

```{r}
seq(-100, 200, by = 0.1) -> my_numbers
str(my_numbers)
```

```{r}
letters
LETTERS
```

## creating data frames from vectors

the most general way to do this is with the `data.frame()`

Remember: vectors must have the same length to be part of a data frame

```{r}
data.frame(names = our_names, numbers = 1:7)
```

Even though we can't use vectors that are larger to create a data frame, we can use the vector that is shorter, and R will recycle the values:

```{r}
data.frame(names = our_names, numbers = 1)
```
R will recycle the values only if they are multiples of the vector

To recycle the values of a numeric vector of length 2, we have to repeat the vector names two times, so it is a
```{r}
data.frame(names = rep(our_names,2), numbers = c(2, 5.5))
```

## Exercise 6
```{r}
length <- c(2.2, 2.1, 2.7, 3.0, 3.1, 2.5, 1.9, 1.1, 3.5, 2.9)
width <- c(1.3, 2.2, 1.5, 4.5, 3.1, NA, 1.8, 0.5, 2.0, 2.7)
height <- c(9.6, 7.6, 2.2, 1.5, 4.0, 3.0, 4.5, 2.3, 7.5, 3.2)
data.frame(length = length ,width = width , height = height, genus = "taxus", species = "baccata")


```
# Extracting/accessing data values from vectors and data frames 

```{r}
surveys <- read.csv(file = "../data-raw/surveys.csv")
str(surveys)
```

One common way to extract or access vectors from column in a data frame is the dollar sign symbol `$`
```{r}
surveys$record_id -> record_id
```

Another way is with the square brackets `[]`

```{r}
surveys[1:10, "hindfoot_length"]
```

If I want all of the values from the rows of column hindfoot_length:

```{r}
surveys[,"hindfoot_length"] -> hindfoot_length
```

Another way is using double square brackets 
```{r}
surveys[["record_id"]] -> surveys_record_id
head(surveys_record_id)
```

## Exercise 7
```{r}
surveys$weight -> surveys_weight
surveys[["month"]] -> surveys_month 
surveys[["hindfoot_length"]] -> surveys_hindfoot_length 
surveys$hindfoot_length -> hindfoot_length
na.omit(hindfoot_length)
mean(hindfoot_length)



```

Homework
# Exercise 8




